maxent model avec xvalidation (year-based test training)

### 2019-05-22 src/asc_file_database.Rmd
minor adjustment

```{r pre-install packages, message=FALSE}
packages <- c("tidyr", "rJava", "dismo")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r base list or ascii layers}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85", "1151.99", "1265.86", "1387.38") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

>>>>

```{r pre-install packages, message=FALSE}
packages <- c("tidyr")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

```{r base list or ascii layers}
ascyears <- as.character(c("1998", "1999", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")) 
#unyears <- paste0("X", unyears) #this adds an X to each year to match with raster naming
ascmonths <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
ascdepths <- c("0.50576", "1.55586", "2.66768", "3.85628" , "5.14036", "6.54303", "8.09252", "9.82275", "11.7737" , "13.991", "16.5253" , "19.4298", "22.7576", "26.5583", "30.8746", "35.7402", "041.18", "47.2119", "53.8506", "61.1128" , "69.0217" , "77.6112", "86.9294", "97.0413", "108.03", "000120", "133.076", "147.406", "163.165" , "180.55", "199.79", "221.141", "244.891" , "271.356", "300.888", "333.863" , "370.689", "411.794", "457.626", "508.64", "565.292", "628.026", "697.259", "773.368", "856.679", "947.448", "1045.85") 
#this depth list is based on how the files are named when split into depth layers. note that chl and o2 have had two minor name changed - 163.164 to 163.165 and 370.658 to 370.659, to match all layers

### 2019-05-22 src/maxent_avec_xvalidation.Rmd
minor adjustment

select month
```{r month of interest -change}
mthna <- "oct"
mthno <- "10"
```

@@

```{r binary map per year 2D representation}
binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
colnames(yrlayer) <- c("layer", "year")

@@

```{r binary map averaged 2D representation}
binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*.asc", full.names = TRUE)

@@

```{r calc no presence cells per layer (avg)} 
ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*asc", full.names = TRUE)
binaryavgstack <- stack(ascdepth_list)
stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstat[ ,1] <- c(1:49)
colnames(stackstat) <- c("layer", "cell_sum")

@@

@@

```{r graph by depth (avg)}
bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1))
trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
print(bydepth)
dev.off()
```

@@

```{r high probability map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
colnames(yrlayer) <- c("layer", "year")

@@

```{r high probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))
dev.off()
```
```{r med probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_medavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_avg_2d.png")))
dev.off()
```

```{r low probability average map per year 2D representation}
binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_lowavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_avg_2d.png")))
dev.off()
```

@@

binaryavgstack <- stack(ascdepth_list_high)
stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstathigh[ ,1] <- c(1:49)
colnames(stackstathigh) <- c("layer", "high_cell_sum")

@@

binaryavgstack <- stack(ascdepth_list_med)
stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatmed[ ,1] <- c(1:49)
colnames(stackstatmed) <- c("layer", "med_cell_sum")

@@

binaryavgstack <- stack(ascdepth_list_low)
stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatlow[ ,1] <- c(1:49)
colnames(stackstatlow) <- c("layer", "low_cell_sum")

@@

write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)
```

@@

```{r graph by depth - prob plot (avg)}
bydepth <- ggplot(data = stackstat, aes(x = layer, y = value, fill = variable)) + geom_bar(stat = "identity") +   scale_x_reverse() +
coord_flip()
png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
print(bydepth)
dev.off()
```

>>>>

select month
```{r month of interest -change}
mthna <- "jan"
mthno <- "1"
```

@@

```{r binary map per year 2D representation}

binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryclog_list)))
colnames(yrlayer) <- c("layer", "year")

@@

```{r binary map averaged 2D representation}

binaryclog_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*.asc", full.names = TRUE)

@@

```{r calc no presence cells per layer (avg)} 
ascdepth_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_cloglog.*asc", full.names = TRUE)
binaryavgstack <- stack(ascdepth_list)
stackstat <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstat[ ,1] <- c(1:46)
colnames(stackstat) <- c("layer", "cell_sum")

@@

attach depth in meters (layer_name in depth_bin.csv) to the stackstat based on layer (stackstat) and depthlayerno (depth_bin.csv)

```{r}
dep <- read.csv("../data/env/depth_bin.csv", header = TRUE) # the depth bin file created way back
dep <- subset(dep, select = c("depthlayerno", "layer_name"))
stackstat <-  read.csv(paste0("../output/maxent/predictions/", mthna, "/cellcount_bydepth_avg.csv"), header = TRUE)
colnames(dep)[names(dep) == "depthlayerno"] <- "layer"
stackstat <- left_join(stackstat, dep)
stackstat[ , "layer_name"]=round(stackstat[ , "layer_name"], 2)
head(stackstat)

@@

```{r graph by depth (avg)}
bydepth <- barchart(layer ~ cell_sum, data = stackstat, ylim = c(49, 1))
trellis.device(device="png", filename = paste0("../output/maxent/predictions/", mthna, "/depthplot_avg.png"))
print(bydepth)
dev.off()

@@

```{r high probability map per year 2D representation}

binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_highprob_cloglog.*.asc", full.names = TRUE)
yrlayer <- data.frame(matrix(ncol = 2, nrow = length(binaryprob_list)))
colnames(yrlayer) <- c("layer", "year")

@@

```{r high probability average map per year 2D representation}

binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_highprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_highavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_highprob_avg_2d.png")))
dev.off()

```{r med probability average map per year 2D representation}

binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_medprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_medavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_medprob_avg_2d.png")))
dev.off()

```{r low probability average map per year 2D representation}

binaryprob_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^binary_avg_lowprob.*.asc", full.names = TRUE)

binstack <- stack(binaryprob_list)
stacksum <- sum(binstack, na.rm = TRUE)
writeRaster(stacksum, paste0("../output/maxent/predictions/", mthna, "/binary_lowavg_2d.asc"), overwrite=TRUE)
plot(stacksum, png(paste0("../output/maxent/predictions/", mthna, "/binary_lowprob_avg_2d.png")))
dev.off()

@@

binaryavgstack <- stack(ascdepth_list_high)
stackstathigh <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstathigh[ ,1] <- c(1:46)
colnames(stackstathigh) <- c("layer", "high_cell_sum")

@@

binaryavgstack <- stack(ascdepth_list_med)
stackstatmed <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatmed[ ,1] <- c(1:46)
colnames(stackstatmed) <- c("layer", "med_cell_sum")

@@

binaryavgstack <- stack(ascdepth_list_low)
stackstatlow <- data.frame(matrix(ncol = 2, nrow = nlayers(binaryavgstack)))
stackstatlow[ ,1] <- c(1:46)
colnames(stackstatlow) <- c("layer", "low_cell_sum")

@@

write.csv(stackstat, paste0("../output/maxent/predictions/", mthna, "/probability_cellcount_bydepth_avg.csv"), row.names = FALSE)

@@

```{r graph by depth - prob plot (avg)}
bydepth <- ggplot(data = stackstat, aes(x = layer, y = value, fill = variable)) + geom_bar(stat = "identity") +   scale_x_reverse() +
coord_flip()
png(paste0("../output/maxent/predictions/", mthna, "/depthplot_probavg.png"))
print(bydepth)
dev.off()
